Greenhouse gas emissions model

### 2021-07-13 210617_GHG_model.Rmd
clean and add citaitons/ uncertainty for Ch4 and milk calculations (placeholders for now)

```{r, warning = F, include = F}
#install and load packages
library(decisionSupport)
library(rmarkdown)

@@

```{r}

library(decisionSupport)
###CW Note ####
# remove all the 'C:/Users/jwafula/Dropbox/UNIQUE_CCAFS/analysis/data/' business
###CW Note ####
### # This project will be stored in relative file paths (i.e. same or lower folders)
input_table <- "livestock_ghg_input_table.csv"

# Generate a function for testing the model function 'line by line' 
# by taking a single random sample of the provided estimates
make_variables <- function(est,n=1){ x <- random(rho=est, n=n)
for(i in colnames(x)) assign(i, as.numeric(x[1,i]),envir=.GlobalEnv)}
# run the function on the input_table
make_variables(estimate_read_csv(input_table))

# Generate the GHG emissions function 
ghg_emissions<-function(x, varnames){

# Enteric fermentation emissions ($CH_4$)
  ####CW Note ####
  #in these code sections I am missing the overview. 
  # it is not immediately clear what this is meant to do... 
  # from a specific publication?

@@

#Net energy for growth (NE_growth; MJ/day):

NE_growth<-4.18*((0.0635*478/(C*FSBW))^0.75*(0.96*SWG)^1.097)

# here, C is an animal type coefficient
# FSBW: final shrunk body weight in kg - this is the actual final body weight multiplied by 0.96 (to subtract the last meal)

@@

NE_DE<-(1.123-(4.092*(10^-3)*dig_energy)+(1.126*(10^-5)*(dig_energy^2))-(25.4/dig_energy))

NEg_DE<-(1.164-(5.160*(10^-3)*dig_energy)+(1.308*(10^-5)*(dig_energy^2))-(37.4/dig_energy))

@@

gross_energy_intake<-(((NE_m+NE_mobilized+NE_activity+NE_lactation+NE_w+NE_pregnancy)/NE_DE)+(NE_growth/NEg_DE))/(dig_energy/100)

@@

##### EL CW Note #### 
#Y_m here is the methane conversion rate - Joshua used the IPCC default value of 6.5%. Full story is on page 302 (a bit more complicated, with differentiation by animal type)

# Compute the CO2 equation for enteric methane
##### EL CW Note #### 
##### #This equation originally included multiplication by 365 - this seems to have been erroneous (already done in the previous line)
#This converts methane emissions to their Global Warming Potential (CO2-equivalent of CH4 is 25)
enteric_CO2=enteric_CH4*25

@@

##### WE STOPPED HERE

# Calculate Nitrous oxide emissions from manure and urine on pasture
n_intake<-(gross_energy_intake/18.45)*((pregnancy_coefficient_Cp/100)/6.25)
n_retension<-((milk_yield*(3.3/100))/6.38)
n_excretion<-n_intake*(1-n_retension)*365
mm_direct_n2o<-n_excretion*((mms_pasture*0.02)+(mms_spread*0.005)+(mms_drylot*0.02)+(mms_solidstore*0.005)+
               (mms_composted*0.006)+(mms_liquid*0.005)+(mms_biogas*0)+(mms_burn*0))*0.01*(44/28)

@@

# plot

@@

return(list(enteric_CH4=enteric_CH4,
            #CW NOTE ####
            #working backwards from 'enteric_CO2' I see that 
            # this is the product of something twice multiplied by 365
            enteric_CO2=enteric_CO2,
            mm_CO2=mm_CO2,
            mm_CO2_N2O=mm_CO2_N2O,
            feed_CO2=feed_CO2,
            on_farm=on_farm))
}

@@

>>>>

```{r, warning = F, include = F}
#install and load packages
library(decisionSupport)
library(ggplot2)
library(plyr)
library(rmarkdown)
# devtools::install_github("CWWhitney/uncertainty")
library(uncertainty)

@@

```{r load_data}

library(decisionSupport)
###CW Note ####
# remove all the 'C:/Users/jwafula/Dropbox/UNIQUE_CCAFS/analysis/data/' business
###CW Note ####
### # This project will be stored in relative file paths (i.e. same or lower folders)
input_table <- "data/livestock_ghg_input_table.csv"
```

```{r make_variables}
# Generate a function for testing the model function 'line by line' 
# by taking a single random sample of the provided estimates
make_variables <- function(est,n=1){ x <- random(rho=est, n=n)
for(i in colnames(x)) assign(i, as.numeric(x[1,i]),envir=.GlobalEnv)}

# run the function on the input_table
make_variables(estimate_read_csv(input_table))
```

```{r ghg_emissions_function}
# Generate the GHG emissions function 
ghg_emissions<-function(x, varnames){

# Enteric fermentation emissions ($CH_4$)
  ####CW Note ####
  #####in these code sections I am missing the overview. 
  ##### it is not immediately clear what this is meant to do... 
  ##### from a specific publication?

@@

#Net energy for growth (NE_growth; MJ/day):
# here, C is an animal type coefficient
# ### CW Note #### Assuming 'C' is C_activity or should it have been animal_type
C <- C_activity
# C <- animal_type
NE_growth<-4.18*((0.0635*478/(C*FSBW))^0.75*
                   (0.96*SWG)^1.097)

# FSBW: final shrunk body weight in kg - this is the actual final body weight multiplied by 0.96 (to subtract the last meal)

@@

NE_DE<-(1.123-(4.092*(10^-3)*dig_energy)+
            (1.126*(10^-5)*(dig_energy^2))-
              (25.4/dig_energy))

NEg_DE<-(1.164-(5.160*(10^-3)*dig_energy)+
            (1.308*(10^-5)*(dig_energy^2))-
              (37.4/dig_energy))

@@

gross_energy_intake<-(((NE_m+NE_mobilized+NE_activity+NE_lactation+NE_w+NE_pregnancy)/NE_DE)+
                          (NE_growth/NEg_DE))/(dig_energy/100)

@@

##### EL CW Note #### 
######Y_m  here is the methane conversion rate -
###### Joshua used the IPCC default value of 6.5%. Full story is on page 302 (a bit more complicated, with differentiation by animal type)

# Compute the CO2 equation for enteric methane
##### EL CW Note #### 
##### #This equation originally included multiplication by 365 - this seems to have been erroneous (already done in the previous line)
#This converts methane emissions to their Global Warming Potential (CO2-equivalent of CH4 is 25)
enteric_CO2<-enteric_CH4*25

@@

##### #####
##### 
##### WE STOPPED HERE
#####
##### #####

# Calculate Nitrous oxide emissions from manure and urine on pasture
# ### ##### CW Note #### 
# ### ##### To get this to run I am assuming a temp 10% pregnancy_coefficient_Cp
pregnancy_coefficient_Cp <- 0.1
### #####
### ####
n_intake<-(gross_energy_intake/18.45)*((pregnancy_coefficient_Cp/100)/6.25)
n_retension<-((milk_yield*(3.3/100))/6.38)
n_excretion<-n_intake*(1-n_retension)*365
mm_direct_n2o<-n_excretion*((mms_pasture*0.02)+(mms_spread*0.005)+(mms_drylot*0.02)+(mms_solidstore*0.005)+
               (mms_composted*0.006)+(mms_liquid*0.005)+(mms_biogas*0)+(mms_burn*0))*0.01*(44/28)

@@

# ### ##### CW Note
# Removing this plot
# plot

@@

return(list(enteric_CH4=enteric_CH4,
            #CW NOTE ####
            # ####working backwards from 'enteric_CO2' I see that 
            # #### this is the product of something twice multiplied by 365
            # ####### 
            enteric_CO2=enteric_CO2,
            mm_CO2=mm_CO2,
            mm_CO2_N2O=mm_CO2_N2O,
            feed_CO2=feed_CO2,
            on_farm=on_farm))
}

@@

Apply the `decisionSupport` tools for running the Monte Carlo [@R-decisionSupport]. See the vignette guides on *applying the mcSimulation function* [@fernandez_applying_2021] and *simulating the cost effectiveness of controlled burning* [@whitney_simulating_2017] for examples.

```{r monte_carlo_simulation}

GHG_simulation <- mcSimulation(
  estimate = estimate_read_csv("data/livestock_ghg_input_table.csv"),
  model_function = ghg_emissions,
  numberOfModelRuns = 1e2, #100 for now
  functionSyntax = "plainNames"
)

```{r plot_distributions}

plot_distributions(mcSimulation_object = GHG_simulation,
                   vars = "enteric_CH4",
                   method = "boxplot_density",
                   old_names = "enteric_CH4",
                   new_names = "Methane Emissions from Enteric Fermentation")

# uncertainty

Here we apply a collection of functions for use in assessing uncertainty from the `uncertainty` package. Install using `devtools::install_github("CWWhitney/uncertainty")`.

Here we use the `varscatter()` function to create a scatter plot of an influence variable of interest {in_var} (x) and an outcome variable {out_var} (y) with associated and estimated densities with loess smooth linear fits density curves. Plot made with the scatter.hist() function in the plyr() package [@R-plyr] in R [@R-base].

Check the `milk_yield` effect on `enteric_CH4`.

```{r uncertainty_package}
uncertainty::varscatter(in_var = GHG_simulation$x$milk_yield, out_var = GHG_simulation$y$enteric_CH4)
```

Use `varkernel()` for another type of graphical representation of the same relationship. `varkernel()` performs a two-dimensional kernel density estimation for the expected value of the outcome variable {out_var} and the influence variable of interest {in_var} using the `kde2d()` function in the `MASS()` package (Venables and Ripley 2002). The function produces a matrix of the estimated density (z) of the expected value of the outcome variable {out_var} (y) and the influence variable of interest {in_var} (x). It looks slightly different than the scatter plot estimates above because the density function restricts the shape of the kernel to a bivariate normal kernel.

Check the effect of {in_var} `milk_yield` on the {out_var} `enteric_CH4`.

```{r name_in_out_vars}
in_var  <- GHG_simulation$x$milk_yield
out_var <- GHG_simulation$y$enteric_CH4
```

```{r varkernel}
uncertainty::varkernel(in_var = in_var, out_var = out_var)
```

`varkernel()` shows a density surface plot of a given influence variable of interest {in_var} (x) and the outcome variable {out_var} (y). The legend shows the value for the estimated density (z). The plot is made with the `filled.contour()` function of the `graphics()` package.

In `varkernel()` the density (z) over the entire plot integrates to one, and therefore represents the relative probability of an observation (the outcome variable {out_var} along y-axis) given a specific value for a given influence variable of interest {in_var} (along x-axis).

## Estimated outcome variable {out_var} given the expected variable

`varkernelslice()` calculates the estimated outcome variable {out_var} given the expected value of the influence variable of interest {in_var}, based on a slice of 'z' from the Kernel density calculated with `varkernel()`. The `expectedvariable` parameter is set to 30.

```{r varkernelslice}
uncertainty::varkernelslice(in_var = in_var, out_var = out_var, expectedin_var = 30)
```

`varkernelslice()` plots the probabilities (shown along the y-axis) for the expected value of the outcome variable {out_var} (shown along the x-axis). Since this is a cut through the density kernel `varkernel()`, which integrates to 1, the probability values are relative, not absolute measures.

## Influencing variable intervals

`varviolin()` determines different possible influencing variable intervals {in_var} by calculating the optimal interval width for the given variable values. This is done using the `IQR()` function in the `stats()` package, after the Freedman-Diaconis rule (IQR = interquartile range).

$Os = 2 * IQRs / (obs_IQRs)^{(1/3)}$

Where $Os$ is The optimal interval width for our sample, $IQRs$ is the interquartile range for our sample and $obs_IQRs$ is the total number of observations for the interquartile range for our sample.

```{r varviolin}
uncertainty::varviolin(in_var = in_var, out_var = out_var)
```

`varviolin()` shows violin plots of influencing variable {in_var} values (x) and outcome variable {out_var} (y) values, with six different intervals of influencing variable {in_var} values. Plot made with `ggplot2()` [@R-ggplot2].

## Probability of outcome variable given variable

`varkernelslicerange()` shows the influencing variable {in_var} value intervals, optimized interquartile ranges shown in `varviolin()` can be used to select a range to slice from the density kernel `varkernel()` as was done for a single variable value in `varkernelslice()`. Here we set the maximum variable value to 4 and the minimum to 1.

```{r varkernelslicerange}
uncertainty::varkernelslicerange(in_var = in_var, out_var = out_var, 
                                 min_in_var = 1, max_in_var = 4)
```

`varkernelslicerange()` plots the probabilities (shown along the y-axis) for the expected value of the outcome variable {out_var} (shown along the x-axis). As with `varkernelslice()` the probability values shown are relative, not absolute measures. They are the result of cuts through the density kernel `varkernel()`, which integrates to 1.
